\documentclass[../Thesis-IJspeert.tex]{subfiles}

\begin{document}

{
\cleardoublepage
\chapter*{Acknowledgements}
\thispagestyle{empty}
Thank you all!

\iffalse
There are many people without whom it would not have been possible to complete this work.  First and foremost I must thank Axel Kuhn, not only for the giving me the chance to undertake this challenge, but for his countless contributions and his guidance along the way.

I am also grateful to everyone I have worked with in the Atom-Photon Connection.  Since I began several people have helped me keep going; Annemarie Holleczek, who first taught me how to survive in a laser lab; Oliver Barter, who somehow kept his good humour when working closely with me for nearly three years; and Dustin Stuart, whose technical expertise with vacuum chambers and atomic physics were almost as important as his calming influence.  I'm also lucky to have shared an office and a group with Marwan Mohammed, who joined with me and continues to surprise me to this day, Klara Theophilo, Naomi Holland and Thomas Doherty.  Thank you all.

I would also like to thank Peter Nisbet-Jones, Jerome Dilley and everybody else whose work before my time laid the foundations for my experiment -- I hope one day my work can provide the same to whomever comes next.

Finally, I have to thank Caitlin Williams for choosing to put up with me, and my family; Mum, Dad and Jenny, who never had that choice.
\fi

%I would like to thank...AK, OB, DS, AH, KT, MM, NH, TD etc.
%
%Allison?
%
%Funding bodies?
%
%Also - the workshop, POTL, Vikas etc.
%
%Without whom I would have one less thesis and several fewer fingers.
%
%Also friends/family.
}
	
\end{document}
